---
title: "Statical Computing Final Project"
author: "Hatice Nurel - Melise Kaya"
date: "07-06-2024"
output: html_document
---

# 1. (3p) Data

Please find your original dataset or datasets; and describe your data in the first step.
(3-5 sentences)

```{r}
# Load dataset and print the first few rows
# Assuming 'data.csv' is the file name of your dataset
data <- read.csv("data.csv")
head(data)
```

# Describe your data here...

# 2. (3p) Exploratory and descriptive data analysis

Use "Exploratory and descriptive data analysis". Talk about your categorical and
quantitative data or your ordinal variables etc. Write down your comments. (3-4 sentences)

```{r}
# Perform exploratory data analysis here...
```
# Your comments...

# 3. (4p) Data Visualization

Use 2 useful, meaningful and different "data visualization techniques" which will help
you understand your data further (distribution, outliers, variability, etc). And use
another 2 visualizations to compare two groups (like female/male; smoker/non-smoker etc).
For each visualization, write down your comments. (2-3 sentences each)

```{r, fig.cap="Visualization 1"}
# First visualization code
```
# Comments for Visualization 1...

```{r, fig.cap="Visualization 2"}
# Second visualization code
```
# Comments for Visualization 2...

# 4. (4p) Confidence Intervals

Build '2 Confidence Intervals' step by step: a) Calculate the mean, then b) standard error,
and then c) the CI. Make "clear comments" about your findings in a, b and c.

```{r}
# Confidence Interval calculation code...
```
# Your clear comments...

# 5. (4p) Transformation

Implement one transformation (log transformation, Box-Cox transformation, etc) for
one of your quantitative variables, which is not normally distributed; but will be
normal or more normal, after the transformation. Comment about your trial (3-4 sentences)

```{r}
# Transformation code...
```
# Your comments...

# 6. (2p every item if not indicated) t-test (Welch t-test or Wilcoxon rank-sum test or Paired t-test)

Implement a single t-test for one of your "normally or not-normally distributed" variable:

## a. Aim
In words, what is your objective here? (1-2 sentences)

# Your aim...

## b. Hypothesis and level of significance
Write your hypothesis in scientific form and determine the level of significance.

```{r}
# Hypothesis formulation...
```

## c. Which test you choose
Is your data independent or dependent? (1 sentence)
Tell why you chose this test. (1 sentence)

# Answer c...

## d. Assumption Check:
Check the required assumptions statistically and "comment (1 sentence) for each of them is a must!".

```{r}
# Assumption check code...
```

# Comments on assumption checks...

## e. Result
Give the output of the test and then write down the result (ex: since p value is less /greater
than alpha, I reject/not reject the null hypothesis). (2-3 sentences)

```{r}
# t-test code...
```
# Your result comment...

## f. Conclusion (4p)
You got your result in item e. Write down the conclusion of your result, in such a way that,
the reader who doesn’t know any statistics can understand your findings. (1-2 sentences)

# Your conclusion...

## g. What can be Type-1 and Type-2 error here? Not definition! Tell these situations in terms of your data. (4p) (2-3 sentences)

# Explanation of Type-1 and Type-2 errors...

# 7. (2p every item if not indicated) ANOVA and Tukey Test

## a. Aim (4p)
In words, what is your objective here? (1-2 sentences)

# Your aim...

## b. Hypothesis and level of significance
Choose more than 2 (≥3) groups to compare! (1 sentence)
Write your hypothesis in scientific form and determine the level of significance.

```{r}
# Hypothesis formulation...
```

## c. Assumption Check
Check the required assumptions statistically. "comment (1 sentence) on each of them is a must!".

```{r}
# Assumption check code...
```

# Comments on assumption checks...

## d. Result of ANOVA
Give the output of the test and then write down the result (ex:since p value is less than alpha,
I reject the null hypothesis) (1-2 sentences)

```{r}
# ANOVA test code...
```
# Your result comment...

## e. Conclusion of ANOVA (4p)
You got your result in item d. Write down the conclusion of your result, in such a way that,
the reader who doesn’t know any statistics can understand your findings. (1-2 sentences)

# Your conclusion...

## f. Result of Tukey
Give the output of the test and write down the result (ex: since p value is less /greater than alpha,
I reject/not reject the null hypothesis) (1-2 sentences)

```{r}
# Tukey test code...
```

# Your result comment...

## g. Conclusion of Tukey (4p)
You got your result in item f. Write down the conclusion of your result, in such a way that, the
reader who doesn’t know any statistics can understand your findings. (2-3 sentences)

# Your conclusion...

# 8. (2p every item) Multiple Linear Regression

## a. Aim (4p)
In words, what is your objective here? Not definition, talk about your own aim/problem. (2-3 sentences)

# Your aim...

## b. Variable Selection
Multiple linear regression (MLR) is a statistical technique that uses several explanatory variables to predict the outcome of a response variable. Which ones are your explanatory variables and which one is your response variable? (2-3 sentences)

# Variables selection explanation...

## c. Regression Equation
Write down the "statistical/mathematical equation" of your regression function using those variables and the parameters.

# Your regression equation...

## d. Hypothesis and level of significance (4p)
Write your hypothesis in scientific form and determine the level of significance.

```{r}
# Hypothesis formulation...
```

## e. Find the Best Model
Use step function and find the best model, describe the reason which makes it the best one.

```{r}
# Model selection code...
```

# Explanation of the best model...

## f. Assumption Check (4p)
Check the required assumptions statistically, "comment (1 sentence) on each of them is a must!".

```{r}
# Assumption check code...
```

# Comments on assumption checks...

## g. Result
Give the output of the best model and write down the result. (2-3 sentences)

```{r}
# Model result code...
```

# Your result comment...

## h. Conclusion (4p)
You got your result in item f. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings. (2-3 sentences)

# Your conclusion...
```
Please note that you need to replace placeholder texts with your actual content, formulas, hypotheses, and R code where appropriate. The above format assumes you've installed R and required packages for your analysis.